STA610 Case Study 1 Team 4 Report

### 2021-03-21 report.Rmd
draft interpretation

---
title: "STA610 Case Study 1 Team 4 Report"
author: 
  - Cole Juracek (Programmer) 
  - Lauren Palazzo (Programmer) 
  - Lingyu Zhou (Checker) 
  - Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, lubridate)
#devtools::install_github("goodekat/redres")
```

@@

print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

@@

## Model

Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone. We are also specifically asked to control for location and how drug price changes with it. Cities seemed to be too noisy to use as a location predictor. On the flip side, region did not seem to be associated with a significant change in drug prices. Therefore, state provides the best granularity of accounting for location in this instance.

Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price. As seen in the EDA, however, many states have excessively small sample sizes. This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance. We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.

This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.

#### Model Specification

$$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$

$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \beta_{5} x_{ij5} + \epsilon_{ij}$$

$$
\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
$$
$$
\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
$$

This linear model estimates the log price per miligram $log(y)$ of oxymorphonefor an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:

* $x_1$: The number of days since the recorded transaction and January 1st, 2000
* $x_2$: Dosage strength in mg of the units purchased
* $x_3$: Bulk purchase, an indicator for whether 10+ units were purchased at once
* $x_4$: The primary reason for purchase. Factors include "other/unknown", "self-treatment", and "enjoyment"
* $x_5$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "heard it", and "personal"

@@

- mgstr: good
- bulk_purchase: good

```{r}
mod1b.inf <- influence(mod1b, maxfun=1)
infIndexPlot(mod1b.inf, vars='Cook')
```

```{r}
#mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
#mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
big_data <- df[!influential,]
mod1c <- lmer(data=big_data, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
qqmath(mod1c)
```

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
# resqq <- plot_resqq(mod1)
# ranefci <- plot_ranef(mod1)
# grid.arrange(resqq, ranefci, ncol = 2)
```

@@

@@

* Bulk_purchase: “Not bulk purchase” is the reference group.
Controlling for other variables, being bulk purchase increases the price per mg by exp(-0.16) = 0.8521. 
* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
Controlling for other variables, taking oxymorphone with the primary reason of self-treating tends to receive a exp(0.05) = 1.0513 times higher price per mg than people whose primary reason is “other or unknown”.
Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to receive a exp(0.02) = 1.0202 times higher price per mg than people whose primary reason is “other or unknown”.
* Source: “Source: Unknown” is the reference group. 
Controlling for other variables, purchasing oxymorphone with Internet source tends to receive a exp(0.02) = 1.0202 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with personal source tends to receive a exp(0.07) = 1.0725 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.

@@

>>>>

---
title: "STA610 Case Study 1 Team 4 Report"
author: 
  - Cole Juracek (Coordinator) 
  - Lauren Palazzo (Programmer) 
  - Lingyu Zhou (Checker) 
  - Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
#devtools::install_github("goodekat/redres")
library(redres)
```

@@

#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

@@

## Model
#### Model Specification

$$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$
$$
\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
$$
$$
\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
$$

* i stands for individuals and j stands for each state. 
* $\beta_{0,j}$ is the random intercept by state. 
* $\beta_{1}$ measures the fixed effect for date_num, the date variable.
* $\beta_{2}$ measures the fixed effect for mgstr, the dosage strength of reported drug in milligrams.
* $\beta_{3}$ measures the fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
* $\beta_{4}$ measures the fixed effect for the categorical variable Primary_Reason which contains three categories: Other or unknown, Self-treat and Enjoyment.
* $\beta_{5}$ measures the fixed effect for the categorical variable source

@@

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
```

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
resqq <- plot_resqq(mod1)
ranefci <- plot_ranef(mod1)
resi <- plot_redres(mod1, type = "std_cond")
grid.arrange(resi, resqq, ranefci, ncol = 3)
```

@@

From the esiduals versus the fitted values plot, there are several outliers in the lower half of the plot. 
From the normal quantile plot of the raw conditional residuals, the residuals are skewed and there are a lot of outliers. Thus, the residuals are not normally distributed. 
From the normal quantile plot on random effect state, there are only two outliers. Thus, we can accept random effect of state is normally distributed.

@@

* date_num: For the date_num variable, controlling for other variables, for every one unit increase in the date_num, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0001290) =  0.999871, about a 0.0129% decrease.

* mgstr: For the mgstr variable, controlling for for other variables, for every one unit increase in mgstr, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0241262) =  0.9761625, about a 2.38375% decrease.

* bulk_purchase: “Not bulk purchase” is the reference group.
  * Controlling for other variables, being bulk purchase reduces the price per mg by a multiplicative effect of exp(-0.161) = 0.8521, about a 14.79% reduction.

* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
  * Controlling for other variables, taking oxymorphone with the primary reason of self-treating increase the price per mg by a multiplicative effect of exp(0.05) = 1.0513, about a 5.13% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”. 
  * Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.2) = 1.221403, about a 22.1403% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.

* source: “Source: Unknown” is the reference group. 
  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.02) = 1.0202, about a 2.02% increase than the price per mg of oxymorphone for people with unknown drug source.
  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.07) = 1.0725, about a 7.25% increase than the price per mg of oxymorphone for people with unknown drug source.
  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.24) = 1.2712, about a 27.12% increase than the price per mg of oxymorphone for people with unknown drug source.

@@

```{r, include=FALSE}
library(sjstats)
icc(mod1)
```

```{r}
ICC <- 0.047
knitr::kable(ICC, col.names = "Intraclass Correlation")
```
Intraclass correlation is 0.047, which means values are not very similar from the same state.
